Measuring Individual Partisan Harm in Redistricting

### 2021-09-12 paper/paper.Rmd
edits

```{r nj-maps, fig.cap="Baseline vote patterns, and two hypothetical gerrymanders. Census-designated places are shaded in gray.", fig.width=6.5, fig.height=3.5}
include_graphics(here("paper/figures/nj_maps.pdf"))
```

@@

The efficiency gap and $F(q)$ both correctly classify the gerrymandered plans as such, and with the correct sign.
Both gerrymanders also have a higher average harm than any of the simulated plans.
In contrast, under the partisan bias measure the Republican gerrymander is completely unbiased.
And according to the mean-median measure both gerrymanders strongly favor Democrats.
For the utility measures, the Democratic plan is a low-utility outlier.
But the Republican plan is near the median for both global and local utility.

How to square this with the finding that the Republican plan is an unfair gerrymander which harms more voters than expected under a neutral process?
As the distribution of expected Democratic seats shows, Democrats are structurally favored in New Jersey, able to win a larger fraction of seats than their statewide vote share under a neutral process.
Plans which actively counteract this to try to draw proportional districts, like the Republican gerrymander, will necessarily harm Democratic voters versus the neutral counterfactual.
But in doing so they will create a more proportional plan that will tend to pack voters of each party into more homogeneous districts, increasing both global and local utility.

>>>>

```{r nj-maps, fig.cap="Baseline vote patterns, and two hypothetical gerrymanders. Census-designated places are shaded in gray and county lines in white.", fig.width=6.5, fig.height=3.5}
include_graphics(here("paper/figures/nj_maps.pdf"))
```

@@

The efficiency gap and $F(q)$ both correctly classify the gerrymandered plans as such, and with the correct sign.
Both gerrymanders also have a higher average harm than any of the simulated plans.
In contrast, under the partisan bias measure the Republican gerrymander is completely unbiased, and according to the mean-median measure, both gerrymanders strongly favor Democrats.

For the utility measures, the Democratic plan is a low-utility outlier.
But the Republican plan is near the median for both global and local utility.
How to square this with the finding that the Republican plan is an unfair gerrymander which harms more voters than expected under a neutral process?
As the distribution of expected Democratic seats shows, Democrats are structurally favored in New Jersey, able to win a larger fraction of seats than their statewide vote share under a neutral process.
Plans which actively counteract this to try to draw proportional districts, like the Republican gerrymander, will necessarily harm Democratic voters versus the neutral counterfactual.
But in doing so they will create a more proportional plan that will tend to pack voters of each party into more homogeneous districts, increasing both global and local utility.
